---
title: "Exploring Maxquant Output With SQLITE"
author: "Witold Wolski"
date: "5 October 2015"
output: html_document
---

# Generate sql database from Maxquant output

[Maxquant](http://www.maxquant.org) produces quite convoluted output in tab delimited files. To get a first idea obout those files I would like to read them into a sql database and explore them using [sqlitestudio](http://sqlitestudio.pl/).

To populate the database I am going to use __R__ and the package _dplyr_.

## Collect meta information about the files


```{r}

rm(list=ls())
suppressMessages(library(dplyr))

#path <- "output/20150930/combined/txt/"

path <- "c:/processing/external/EXTERNAL_0/paolo_20151029_MSQC1/combined/txt"
files <- dir(path)
filesf <- dir(path, full.names = T)


files <- data_frame(files,filesf)

files

info <- file.info(files$filesf)
stopifnot(files$filesf == rownames(info))
files <- tbl_df(data.frame(files, info))
glimpse(files)

```

Select text files with size greater than 0.

```{r}
files <- filter(files, size > 0)
files <- files[ grep("\\.txt",files$files),]

```

## Read all the files into R

```{r}
mqdata <-list()
for(file in files$filesf){
  mqdata[[file]] <- read.csv(file,sep="\t",stringsAsFactors = FALSE)
}
stopifnot(names(mqdata) == files$filesf)

```

## Dump them into an sqlite database

Before doing it you need to fix the database files which we derive from the file names.
As a collegue of mine says if you see a file name such as "Oxidation (M)Sites.txt" you know where they _came from_. It is a lonely place.

```{r, results='hide'}
names <- files$files
names <- gsub("\\.txt", "",files$files)
names <- gsub("[[:space:]]", "", names) 
names <- gsub("[[:punct:]]", "", names)
names(mqdata) <- names

library(DBI)
if(file.exists(file.path(path,"my_mqdb.sqlite3"))){
  file.remove(file.path(path,"my_mqdb.sqlite3"))
}
  
my_mqdb <- src_sqlite(file.path(path,"my_mqdb.sqlite3"), create=TRUE)

for( i in 1:length(mqdata)){
  tmp<-copy_to(my_mqdb, mqdata[[i]], name=names(mqdata)[i], temporary = FALSE)
}
dbDisconnect(my_mqdb$con)

```

Now you are ready to explore the data with SQLiteStudio.

For more interesting information about Data Science in Mass Spectrometry visit [http://aggrivet.blogspot.ch/](http://aggrivet.blogspot.ch/).
