---
title: "Join Tables On Float Values With Errors"
author: "Witold Wolski"
date: "6 October 2015"
output: 
  html_document: 
    toc: yes
---

# Problem definition

In mass spectrometry you frequently need to match points in RT and MZ with some measurement error.



# Generate sample data

```{r}



tmp1 <- cbind(sample(letters,1000,replace=TRUE),sample(letters,1000,replace=TRUE),sample(letters,1000,replace=TRUE))
tmp2 <- cbind(sample(letters,1000,replace=TRUE),sample(letters,1000,replace=TRUE),sample(letters,1000,replace=TRUE))

tmp1<-apply(tmp1, 1, paste ,collapse="")
tmp2<-apply(tmp2, 1, paste ,collapse="")

nrseq1 <- 1:1000
nrseq2 <- (rnorm(1000, mean=1000,sd=100))

tab1 <- data.frame(tmp1, nrseq1 + rnorm(1000,0,1.5), nrseq2 + rnorm(1000,0,0.4))
tab2 <-  data.frame(tmp2,  nrseq1 + rnorm(1000,0,1.5), nrseq2 + rnorm(1000,0,0.4) )

colnames(tab1) <- c("name","rt","mz")

tab2 <-  data.frame(tmp2,  nrseq1 + rnorm(1000,0,1.5), nrseq2 + rnorm(1000,0,0.4) )
colnames(tab2) <- c("name","rt","mz")
```

Add 4 columns with error range for mz and rt to one of the tables

  - mzmi - min mass
  - mzma - max mass
  - rtmi - min rt
  - rtma - max rt


```{r}

mzerror <- 1
rterror <- 2.5
tab2<-data.frame(tab2, rtmi = tab2$rt - rterror , rtma = tab2$rt + rterror , mzmi = tab2$mz - mzerror, mzma = tab2$mz + mzerror )

```

# Using DBI and dplyr

## Dump both dataframes into a sqlite database using dplyr

```{r}
suppressMessages(library(dplyr))
suppressMessages(library(DBI))

dbfile <- "my_mqdb.sqlite3"
if(file.exists(dbfile)){
  file.remove(dbfile)
}

my_mqdb <- src_sqlite(dbfile, create=TRUE)
copy_to(my_mqdb, tab1,  temporary = FALSE)
copy_to(my_mqdb, tab2,  temporary = FALSE)


dbDisconnect(my_mqdb$con)

```

## Join both tables on double columns to find matching features

```{r}

my_mqdb <- src_sqlite(dbfile)
dbListTables(my_mqdb$con)
res <-dbSendQuery(my_mqdb$con , "select tab1.mz, tab1.rt, tab2.mz as tab2mz, tab2.rt as tab2rt from tab1 
                  inner join tab2 on tab1.mz < tab2.mzma and tab1.mz > tab2.mzmi 
                  and tab1.rt < tab2.rtma and tab1.rt > tab2.rtmi")
resData <- dbFetch(res,n=-1)
dbClearResult(res)
dim(resData)
head(resData)
plot(resData$rt, resData$mz, xlab="RT", ylab="mz",pch="x" )
points(resData$tab2rt, resData$tab2mz,col=2 )
plot((resData$rt + resData$tab2rt)/2, resData$mz - resData$tab2mz, xlab="rt")
plot(resData$rt - resData$tab2rt, (resData$mz + resData$tab2mz)/2, ylab="mz")
dbDisconnect(my_mqdb$con)

```

AFAIK you can't do it as easily with any of the R join methods as `base::merge` or `dplyr::innter_join`. But maybee sqldf does it?

# Joining using sqldf package

```{r}
suppressMessages(library(sqldf))
res2 <- sqldf("select tab1.mz, tab1.rt, tab2.mz as tab2mz, tab2.rt as tab2rt from tab1 
                  inner join tab2 on tab1.mz < tab2.mzma and tab1.mz > tab2.mzmi 
                  and tab1.rt < tab2.rtma and tab1.rt > tab2.rtmi")
dim(res2)
head(res2)
plot(res2$rt, res2$mz, xlab="RT", ylab="mz",pch="x" )
points(res2$tab2rt, res2$tab2mz,col=2 )


```